A Guide to the Continuous Constant pH Molecular Dynamics Methods in Amber and CHARMM v[\versionnumber]

### 2022-08-09 Old/2_prereq.tex

\section{Prerequisites} % - Ruibin
CpHMD is a specialized MD technique, so the user should have some familiarity with MD simulations and especially with the Amber or CHARMM packages. 
Users with experience with other MD engines, such as Gromacs
\cite{Abraham_Lindahl_2015_SoftwareX}, NAMD\cite{Phillips_Tajkhorshid_2020_J.Chem.Phys.}, 
or OpenMM
\cite{Eastman_Pande_2017_PLoSComput.Biol.} 
are encouraged to go through the introductory tutorials for Amber or CHARMM to get started. 
Before running a CpHMD simulation, it is important to know your research objective. 
If the objective is to predict the {\pka} values of soluble proteins or their protonation states at a certain pH condition, GBNeck2-CpHMD 
\cite{Huang_Shen_2018_J.Chem.Inf.Model.,Harris_Shen_2019_J.Chem.Inf.Model.}
is the best choice and Amber \cite{Case_Kollman_2018_}
experience is required. 
If the objective is to investigate the detailed proton-coupled conformational dynamics, ion transfer, protein-ligand binding/unbinding,
or transmembrane proteins, then 
the hybrid-solvent CpHMD  \cite{Wallace_Shen_2011_J.Chem.TheoryComput.,Huang_Shen_2016_Nat.Commun.} 
or the all-atom PME CpHMD \cite{Huang_Shen_2016_J.Chem.TheoryComput.}
in CHARMM
\cite{Brooks_Karplus_2009_J.Comput.Chem.}
should be used.
We note, a GPU-accelerated all-atom PME CpHMD 
implementation in Amber is forthcoming (Harris, Liu, and Shen, to be published). 
Like all computational chemistry calculations, it is important to know your system as much as possible before any serious simulations. 
With such knowledge, we can decide what salt concentration, temperature, pH range, and titratable residue types to use in CpHMD. Also, the research objective dictates the simulation length. For example, to predict protonation states, simulation can be run until the unprotonated fractions at all pH are converged, typically 10-50 ns per replica. If however, a large conformational transition is of interest, the simulation should be much longer.

@@

 The next step is to add capping ligands ACE and NH2 to the N- and C-terminus and to modify ASP, GLU and HIS to AS2, GL2, and HIP residues if these residue types are set as titratable. For AS2 and GL2, a library file \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-prep/-/blob/master/Files/phmd.lib}{phmd.lib} and a force field modification file \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-prep/-/blob/master/Files/frcmod.phmd}{frcmod.phmd} are included in Ambertools after version 19 \cite{Case_Kollman_2018_}, or we can load them to tLeap/Leap from the directory containing the two files. The last step is to use tLeap/Leap in Amber/ Ambertools to generate topology and parameter files. Our \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-prep}{cphmd-prep} tool set uses the pdbfixer tool for the first step and can do all the aforementioned steps to prepare GBNeck2-CpHMD compatible input files.

>>>>

\section{Prerequisites} % - Ruibin
CpHMD is a specialized MD technique, so the user should have some familiarity with MD simulations and especially with the Amber or CHARMM packages. 
Users with experience with other MD engines, such as Gromacs
\cite{Abraham_Lindahl_2015_SoftwareX}, NAMD\cite{Phillips_Tajkhorshid_2020_J.Chem.Phys.}, 
or OpenMM
\cite{Eastman_Pande_2017_PLoSComput.Biol.} 
are encouraged to go through the introductory tutorials for Amber or CHARMM to get started. 
Before running a CpHMD simulation, it is important to know your research objective. 
If the objective is to predict the {\pka} values of soluble proteins or their protonation states at a certain pH condition, GBNeck2-CpHMD 
\cite{Huang_Shen_2018_J.Chem.Inf.Model.,Harris_Shen_2019_J.Chem.Inf.Model.}
is the best choice and Amber \cite{Case_Kollman_2018}
experience is required. 
If the objective is to investigate the detailed proton-coupled conformational dynamics, ion transfer, protein-ligand binding/unbinding,
or transmembrane proteins, then 
the hybrid-solvent CpHMD  \cite{Wallace_Shen_2011_J.Chem.TheoryComput.,Huang_Shen_2016_Nat.Commun.} 
or the all-atom PME CpHMD \cite{Huang_Shen_2016_J.Chem.TheoryComput.}
in CHARMM
\cite{Brooks_Karplus_2009_J.Comput.Chem.}
should be used.
We note, a GPU-accelerated all-atom PME CpHMD 
implementation in Amber22 \cite{Case_Kollman_2022}
has been recently released \cite{Harris_Shen_Submitted}.
Like all computational chemistry calculations, it is important to know the system of interest
before attempting any simulations. 
With such knowledge, we can decide what salt concentration, temperature, pH range, and titratable residue types to use in CpHMD. Also, the research objective dictates the simulation length. For example, to predict protonation states, simulation can be run until the unprotonated fractions at all pH are converged, typically 10-50 ns per replica. If however, a large conformational transition is of interest, the simulation should be much longer.

@@

 The next step is to add capping ligands ACE and NH2 to the N- and C-terminus and to modify ASP, GLU and HIS to AS2, GL2, and HIP residues if these residue types are set as titratable. For AS2 and GL2, a library file \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-prep/-/blob/master/Files/phmd.lib}{phmd.lib} and a force field modification file \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-prep/-/blob/master/Files/frcmod.phmd}{frcmod.phmd} are included in Ambertools after version 19 \cite{Case_Kollman_2018}, or we can load them to tLeap/Leap from the directory containing the two files. The last step is to use tLeap/Leap in Amber/ Ambertools to generate topology and parameter files. Our \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-prep}{cphmd-prep} tool set uses the pdbfixer tool for the first step and can do all the aforementioned steps to prepare GBNeck2-CpHMD compatible input files.

### 2022-08-09 Old/5_gphmd.tex

\paragraph{Energy minimization}
The energy minimization is performed
with the input file ending with `\_mini.mdin',
e.g., 
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/blob/main/gphmd_Amber/gphmd_snase_phrex/3BDC/3BDC_chainA_A_mini.mdin}{3BDC\_chainA\_A\_mini.mdin} for SNase.
During minimization, a harmonic restraint with the
force constant of 50 kcal mol$^{-1}$ \AA$^{-2}$ is placed
on all protein heavy atoms, except for the 
terminal and capping residues where all atoms are allowed to move. 
The minimization lasts 1000 steps, using the steepest descent and 
conjugate gradient algorithms for the first and second 500 steps, respectively. 
To run minimization, we call \textbf{pmemd.MPI} or \textbf{pmemd.cuda} if Amber 20 \cite{Case_Kollman_2021_Amber20} or a later version is used.
\begin{lstlisting}
$ export name='3BDC_chainA_A'
$ mpirun -n 4 
  $AMBERHOME/bin/pmemd.MPI -O 
  -i ${name}_mini.mdin -c ${name}.rst7 
  -p ${name}.parm7 -ref ${name}.rst7 
  -r ${name}_mini.rst7 -o ${name}_mini.out
\end{lstlisting}
% The latter requires at least an Nvidia GPU installed and configured properly as discusses above.
Here, '-n 4' specifies 4 CPU cores. Note, a large number of processors may be specified if hardware allows. 
The `-ref' option 
specifies the reference coordinate file
(3BDC\_ModelA\_A\_fixed.rst7).
The '-l' option specifies the initial coordinate file. It is omitted here, since it the same as the reference file.
If minimization runs error free, a timing info is given out
and two files are generated: 
the file 3BDC\_ModelA\_A\_fixed\_mini.out 
contains the printed energies, and 
*\_mini.rst7 is the restart coordinate file for the next step equilibration.

@@

\paragraph{Production}
For production runs, we can use either independent pH, 
synchronous pH replica exchange, or asynchronous pH replica exchange.
%(\href{https://gitlab.com/shenlab-amber-cphmd/async_ph_replica_exchange}{Async-pHREX}). 
For independent pH runs, the simulation at different pH 
conditions are run independently either on the same or different machines. 
The running order is arbitrary, and we collect all the output files together for data analysis. 
The input files can be generated using 
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-prep/-/tree/master/cphmd_tools/cphmd_prep.sh}{cphmd\_prep.sh}
with \textbf{-run ind}.
For SNase, the production run 
is invoked with the following command:
\begin{lstlisting}
$ export name='3BDC_chainA_A'
$ $AMBERHOME/bin/pmemd.cuda -O
  -i ${name}_pH7.5_prod.mdin
  -c ${name}_equil4.rst7 -p ${name}.parm7
  -r ${name}_pH7.5_prod1.rst7
  -o ${name}_pH7.5_prod1.out
  -x ${name}_pH7.5_prod1.nc
  -phmdin ${name}.phmdin
  -phmdparm gbneck2_input.parm
  -phmdstrt ${name}_equil4.phmdrst
  -phmdout ${name}_pH7.5_prod1.lambda
  -phmdrestrt ${name}_pH7.5_prod1.phmdrst
  -inf pH7.5.mdinfo
\end{lstlisting}
The restart files generated by equilibration step 4 are used as starting files for the production run. No reference file is needed. 
Along with *.out, *.rst7, *.nc (trajectory), 
and *.phmdrst files, 
$\lambda$ coordinates are saved in 3BDC\_chainA\_A\_fixed\_pH*\_prod1.lambda. 
For the command above, we again use pmemd.cuda
but we can also run on CPUs with pmemd.MPI.
All input files for this example can be found in \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/gphmd_Amber/gphmd_snase_phind}{gphmd\_snase\_phind}.

>>>>

\paragraph{Energy minimization}
The energy minimization is performed
with the input file ending with `\_mini.mdin',
e.g., 
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/blob/main/gphmd_Amber/gphmd_snase_phrex/3BDC/3BDC_chainA_A_mini.mdin}{3BDC\_chainA\_A\_mini.mdin} for SNase.
During minimization, a harmonic restraint with the
force constant of 50 kcal mol$^{-1}$ \AA$^{-2}$ is placed
on all protein heavy atoms, except for the 
terminal and capping residues where all atoms are allowed to move. 
The minimization lasts 1000 steps, using the steepest descent and 
conjugate gradient algorithms for the first and second 500 steps, respectively. 
To run minimization, we call \textbf{pmemd.MPI} or \textbf{pmemd.cuda} if Amber 20 \cite{Case_Kollman_2020} or a later version is used.
\begin{lstlisting}
$ export name='3BDC_chainA_A'
$ mpirun -n 4 
  $AMBERHOME/bin/pmemd.MPI -O 
  -i ${name}_mini.mdin -c ${name}.rst7 
  -p ${name}.parm7 -ref ${name}.rst7 
  -r ${name}_mini.rst7 -o ${name}_mini.out
\end{lstlisting}
% The latter requires at least an Nvidia GPU installed and configured properly as discusses above.
Here, '-n 4' specifies 4 CPU cores. Note, a large number of processors may be specified if hardware allows. 
The `-ref' option 
specifies the reference coordinate file
(3BDC\_ModelA\_A\_fixed.rst7).
The '-l' option specifies the initial coordinate file. It is omitted here, since it the same as the reference file.
If minimization runs error free, a timing info is given out
and two files are generated: 
the file 3BDC\_ModelA\_A\_fixed\_mini.out 
contains the printed energies, and 
*\_mini.rst7 is the restart coordinate file for the next step equilibration.

@@

\paragraph{Production}
For production runs, we can use either independent pH, 
synchronous pH replica exchange, or asynchronous pH replica exchange.
%(\href{https://gitlab.com/shenlab-amber-cphmd/async_ph_replica_exchange}{Async-pHREX}). 
For independent pH runs, the simulation at different pH 
conditions are run independently either on the same or different machines. 
The running order is arbitrary, and we collect all the output files together for data analysis. 
The input files can be generated using 
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-prep/-/tree/master/cphmd_tools/cphmd_prep.sh}{cphmd\_prep.sh}
with \textbf{-run ind}.
For SNase, the production run 
is invoked with the following command:
\begin{lstlisting}
$ export name='3BDC_chainA_A'
$ $AMBERHOME/bin/pmemd.cuda -O
  -i ${name}_pH7.50_prod.mdin
  -c ${name}_equil4.rst7 -p ${name}.parm7
  -r ${name}_pH7.50_prod1.rst7
  -o ${name}_pH7.50_prod1.out
  -x ${name}_pH7.50_prod1.nc
  -phmdin ${name}.phmdin
  -phmdparm gbneck2_input.parm
  -phmdstrt ${name}_equil4.phmdrst
  -phmdout ${name}_pH7.50_prod1.lambda
  -phmdrestrt ${name}_pH7.5_prod1.phmdrst
  -inf pH7.50.mdinfo
\end{lstlisting}
The restart files generated by equilibration step 4 are used as starting files for the production run. No reference file is needed. 
Along with *.out, *.rst7, *.nc (trajectory), 
and *.phmdrst files, 
$\lambda$ coordinates are saved in 3BDC\_chainA\_A\_fixed\_pH*\_prod1.lambda. 
For the command above, we again use pmemd.cuda
but we can also run on CPUs with pmemd.MPI.
All input files for this example can be found in \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/gphmd_Amber/gphmd_snase_phind}{gphmd\_snase\_phind}.
